
```{r libraries}

library(tidyverse)
library(ggplot2)
library(zoo)
library(forecast)
library(xts)
library(readxl)

```

### Basic Cleaning & Processing

```{r test}

# Reading the GDP data
RGDP_Data <- read_excel("RGDP Data.xlsx")

# extracting the most revised values/recent data (2024 Q1) 
latest_data <- RGDP_Data$ROUTPUT24Q1

# creating a lag for all quarters
lag(latest_data)

# check is a dataset to validate whether the data is stationary 
check <- data.frame(RGDP_Data$DATE, log(latest_data), lag(log(latest_data)))

check <- check %>% 
  rename(c("Date" = "RGDP_Data.DATE",
           "Raw Data" = "log.latest_data.",
           "First Lag" = "lag.log.latest_data.."))

# calculating growth rate of log(GDP) from one quarter to the next
check <- check[-1,] %>% 
  mutate(growth_rate = (log(`Raw Data`) - log(`First Lag`))/log(`First Lag`) * 100)

# formatting the data variable in terms of year and quarters
Dates <- gsub(":", " ", check$Date) 
check <- check %>% 
  mutate(Time = as.yearqtr(Dates)) %>% 
  select(c(Time, growth_rate)) %>% 
  mutate(growth_rate = as.numeric(growth_rate))

check_xts <- xts(check$growth_rate, check$Time) 

plot(as.zoo(check_xts), 
     plot.type = "single", 
     col = c("darkred"),
     lwd = 1,
     xlab = "Date",
     ylab = "Growth Rate",
     main = "Quarterly Growth Rate of log(GDP)")

# function that transform years to class 'yearqtr'
YToYQTR <- function(years){
  return(
      sort(as.yearqtr(sapply(years, paste, c("Q1", "Q2", "Q3", "Q4"))))
  )
}

# recessions
recessions <- YToYQTR(c(1961:1962, 1970, 1974:1975, 1980:1982, 1990:1991,
                2001, 2007:2008))
# the COVID recession ended in April 2020 according to the Fed
recessions_covid <- append(recessions, 
                           c(as.yearqtr("2020 Q1"), 
                             as.yearqtr("2020 Q2")))
          
# colour shading for recessions
xblocks(time(as.zoo(check_xts)), 
        c(time(check_xts) %in% recessions_covid), 
        col = alpha("steelblue", alpha = 0.3))


## EXPLAIN EVERYTHING 
# variance of initial years (prior to them changing to GDP) seems higher 
```


### AR Function & Model 

```{r autoregression}

# Using the basic AR model from lecture 
# Inputs: Y - predicted variable, p - AR order, h - forecast horizon

fitAR=function(Y,p,h){
  
  # create p lags + forecast horizon shift (=h option)
  aux = embed(Y, p+h)
  
  #  Y variable aligned/adjusted for missing data due to lags
  y = aux[,1] 
  
  # lags of Y (predictors) corresponding to forecast horizon (prevent leakage)
  X = as.matrix(aux[,-c(1:(ncol(Y)*h))])

  # retrieve last p observations
  X.out = tail(aux,1)[1:ncol(X)] 

  # estimate direct h-step AR(p) by OLS 
  model = lm(y~X) 
  
  # extract coefficients
  coef = coef(model) 

  #make a forecast using the last few observations: a direct h-step forecast.
  pred = c(1,X.out)%*%coef 
  
  #note the addition of a constant to the test observation vector
  
  #get unadjusted rmsfe (ignoring estimation uncertainty)
  rmsfe = sqrt(sum(model$residuals^2)/nrow(X)) 
  
  #save estimated AR regression, prediction, and estimated coefficients
  return(list("model"=model,"pred"=pred,"coef"=coef, "rmsfe"=rmsfe)) 
}


# Creating a test dataset that comprises the growth_rate from the check dataset 

test_AR <- as.matrix(check$growth_rate)


fitAR(test_AR, 2, 2)

```


### Preparing for Advanced AR Model 

Our advanced AR model seeks to build on the above basic AR model from the lecture. We are not changing the model itself, but the data that is inputted into this model. The quarterly data undergoes monthly revisions for the first 10 years, leading to initial values being inaccurate. Furthermore, the data is also subject to seasonal revisions (for instance, due to changes in how the calculations are performed), which can lead to sudden jumps in quarterly growth rates. Through processes of transforming the data, we hope to be able to predict the final, revised, values of data based on the initial values provided This would make forecasts more accurate to the true value.


```{r advanced AR prep}

# calculating mean first revision percentage changes 

# extracting initial values of all variables from 1965 Q3 to 2019 Q4 (pre-covid)
# 1965 Q4 onwards is when they first started revising data from the previous quarter 

data_65q3_onwards <- RGDP_Data[c(75:292), c(1:219)]  %>% 
  subset(select = -DATE) 

# convert the above data to a matrix for easier manipulation
m1 <- as.matrix(data_65q3_onwards) 

data_65q3_onwards <- matrix(m1, ncol = ncol(data_65q3_onwards), dimnames = NULL)

# base dataset showing the GDP values across quarters 

GDP_across_q <- RGDP_Data[c(74:293), c(1:220)] %>%
  subset(select = -DATE) %>%
  mutate_all(as.numeric)

colnames(GDP_across_q) <- NULL

GDP_across_q <- as.matrix(GDP_across_q) 

lagged_GDP_across_q <- lag(GDP_across_q)

df <- 100 * (log(GDP_across_q) - log(lagged_GDP_across_q))/log(lagged_GDP_across_q)

# Removing the first and last row since we had added those to compute the lag.
df <- df[-c(1, 220),]

# Revised values is a data frame that shows the growth rate of each quarter as compared to the previous quarter, across revisions. For instance, revision0 is the initial growth rate for each quarter, and revision1 shows the growth rate after values have been revised. 

revised_growth <- data.frame(matrix(NA, nrow = 218, ncol = 0))

for (i in 0:40){
  na_vector = c(replicate(218, NA)) 
  na_vector = as.numeric(na_vector)
  
  values <- diag(df)
  
  col_name = paste0("revision", i)
  
  na_vector[1:(218-i)] = values[1:(218-i)]
  revised_growth[[`col_name`]] = na_vector
  
  df <- df[,-c(1)]
}


# Now, we are looking into how the growth rates change due to revision
lagged_growth <- revised_growth[, -c(ncol(revised_growth))]
revised_growth <- revised_growth[,-c(1)]

# We're taking the average of the change in growth per revision 
change_in_growth <- 100*((revised_growth - lagged_growth)/lagged_growth)

change_in_growth[sapply(change_in_growth, is.infinite)] <- NA
  
change_in_growth <- change_in_growth %>% apply(2, mean, na.rm = TRUE)


```



## Testing 


### Inputs & Basic Manipulation 

```{r input}

example_q = "2014 Q1"
example_yq = as.yearqtr(example_q)

# Formatting dataframe of quarters and their estimated values
working_df = RGDP_Data[, -c(1)] %>%
  mutate_all(as.numeric)

colnames(working_df) <- sapply(seq(as.Date(as.yearqtr("1965 Q4", format = "%Y Q%q")), by="quarter", length.out = ncol(working_df)), function(x) format(as.yearqtr(x), format = "%Y Q%q"))

rownames(working_df) <- sapply(seq(as.Date(as.yearqtr("1947 Q1", format = "%Y Q%q")), by="quarter", length.out = nrow(working_df)), function(x) format(as.yearqtr(x), format = "%Y Q%q"))

# Calculating growth rate of each quarter
lagged_working_df = lag(working_df)
perc_change_df = (100 * (log(working_df) - log(lagged_working_df)) / log(lagged_working_df))[-c(1),]

```


### Basic AR Model 

```{r basic AR}

end_of_row_mark = (example_yq - as.yearqtr("1947 Q1", format = "%Y Q%q")) * 4 - 1

basic_values <- perc_change_df[1:end_of_row_mark,][[`example_q`]]

test <- as.matrix(basic_values)

basic_AR_model = fitAR(test, 2, 2)

basic_example_forecast = basic_AR_model$pred
basic_example_model = basic_AR_model$model
basic_example_forecast

AIC_basicAR = AIC(basic_AR_model$model)
AIC_basicAR


```


### Advanced AR Model 

```{r advanced AR}

# Split dataframe into 10 years before target date and current value of growth for quarters more than 10 years ago
ten_year_mark = (example_yq - 10 - as.yearqtr("1947 Q1", format = "%Y Q%q")) * 4
end_of_row_mark = (example_yq - as.yearqtr("1947 Q1", format = "%Y Q%q")) * 4 - 1

# Most recent values for start of time to 10 years before target date
ancient_values <- perc_change_df[1:ten_year_mark,][[`example_q`]]

# All other values
recent_values <- perc_change_df[(ten_year_mark + 1):
                                  end_of_row_mark,][[`example_q`]]

# Applying approximation of final growth numbers on recent values
forecast_growth = recent_values 
for (i in 1:length(recent_values)){
  for (j in 1:i){
    forecast_growth[i] = forecast_growth[i] * (1 + (change_in_growth[40 - j] / 100) )
  }
}

all_data = c(ancient_values, forecast_growth)


test2 <- as.matrix(all_data)

AR_fitted_model = fitAR(test2, 2, 2)

advanced_example_forecast = AR_fitted_model$pred
advanced_example_model = AR_fitted_model$model
advanced_example_forecast

plot(c(all_data), type = "l", lwd = 1)
plot(c(all_data, advanced_example_forecast), type = "l", lwd = 1)

AIC_advancedAR = AIC(AR_fitted_model$model)
AIC_advancedAR

```

