
```{r import, warning = FALSE, results= "hide", echo = FALSE}

source("GDP Cleaning.R")
source("inputs.R")
source("ADL Data.R")
source("AR_Model_Functions.R")
source("ADL Functions.R")

```


# Poor Outlook Function

This function analyses how many of our ADL predictions until and including the forecast horizon are less than 0. i.e. Predicting a negative GDP Growth Rate. 


```{r poor outlook}

# X variables is a vector; input would be ADL variables
poor_outlook <- function(Y_dataframe, X_variables, f_horizon){
  
  m <- length(X_variables)
  
  forecast_output <- c() 
  for (i in 1:m){
    X_temp <- get(X_variables[i])
    for (j in 1:f_horizon){
      forecast <- (ADL_predict_all(Y_dataframe, X_temp, j))$prediction
      forecast_output <- append(forecast_output, forecast)
    }
  }
  
  # what is the number of forecasts that are less than 0
  num_poor_forecast <- sum(forecast_output < 0)
  
  # what is the percentage of forecasts < 0
  perc_poor_forecast <- num_poor_forecast/length(forecast_output)
  
  # identifying which indicators show a poor forecast
  
  # indexes of the indicators predicting GDP growth rate < 0
  less_than_zero <- which(forecast_output < 0)
  less_than_zero <- ifelse(less_than_zero > 5, less_than_zero %% 5, less_than_zero)
  less_than_zero[less_than_zero == 0] <- 5
  
  # names of unique indicators which predict GDP growth rate < 0
  less_than_zero <- unique(less_than_zero)
  indicators_poor <- X_variables[less_than_zero]
  
  print(unique(indicators_poor))
  print(perc_poor_forecast)
}


poor_outlook(GDPGrowth_ts, ADL_variables, 2)

```


# Abnormalities Function

How many of the indicators have been behaving weirdly in the last 4 quarters? This would indicate an imminent crisis. 

The function returns a list comprising a message and the indicators that are behaving abnormally (if any). If the indicators are behaving normally, the output for indicators will be NULL.


```{r abnormalities}


abnormal <- function(X_variables){
  
  m <- length(X_variables)
  indicators <- c()

  for (i in 1:m){
    X_temp <- get(X_variables[i])
    name_indicator <- X_variables[i]
    
    # median of the X variable
    median_value <- median(X_temp)
    # calculate the median absolute deviation (MAD)
    mad_value <- mad(X_temp)
    
    # set the threshold for outlier detection (3 times the MAD)
    threshold <- 3 * mad_value
    
    # calculate the absolute deviations from the median
    absolute_deviations <- abs(X_temp - median_value)
    
    # identify outliers 
    outliers <- X_temp[absolute_deviations > threshold]
    
    # subset last 4 entries
    last_4_entries <- tail(X_temp, 4)
    
    # check if any of the outliers correspond to the last 4 entries
    if(any(last_4_entries %in% outliers)){
      indicators <- append(indicators, name_indicator)
    }
  }
  
  if (length(indicators) == 0){
    message = "All good!"
  } 
  else {
    "Markets have been deviating from the norm severely in the last year. If a recession is not already taking place, one might occur soon."
  }
  
  return(list("message" = message, "indicators" = indicators))
}

abnormal(ADL_variables)


```


# Aggregate Forecast Function

Aggregates the forecasts

```{r aggregate forecast}




```



